Exciton-polariton Josephson junctions at finite temperatures

### 2017-04-15 writing/phase_transition/phase_transition.tex
srep

\documentclass[aps, pre, preprint, groupedaddress, superscriptaddress, showkeys, showpacs] {revtex4-1}

@@

% Remove ugly line above reverences
\def\bibsection{\section*{\refname}}

@@

@@

\title{{Exciton-polariton Josephson junctions at finite temperatures}}

\author{M. E. Lebedev}
\affiliation{ITMO University, St. Petersburg 197101, Russia}

\author{D. A. Dolinina}
\affiliation{ITMO University, St. Petersburg 197101, Russia}

\author{S. M. Arakelian}
\affiliation{Vladimir State University named after A. G. and N. G. Stoletovs, Gorkii Street 87, Vladimir, Russia}

\author{Tien-Chang Lu}
\affiliation{Department of Photonics, National Chiao Tung University, Hsinchu 300, Taiwan}

\author{A. V. Kavokin}
\affiliation{Spin Optics Laboratory, St. Petersburg State University, Ul’anovskaya, Peterhof, St. Petersburg 198504, Russia}
\affiliation{School of Physics and Astronomy, University of Southampton, SO17 1BJ Southampton, United Kingdom}
\affiliation{Istituto CNR-SPIN, Viale del Politecnico 1, I-00133, Rome, Italy}

\author{A. P. Alodjants}
\email{alexander\_ap@list.ru}
\affiliation{ITMO University, St. Petersburg 197101, Russia}
\affiliation{Vladimir State University named after A. G. and N. G. Stoletovs, Gorkii Street 87, Vladimir, Russia}

\date{\today}

\begin{abstract}
We consider finite temperature effects in a non-standard Bose-Hubbard model for an exciton polariton Josephson junction (EPJJ) that is characterised by complicated potential energy landscapes (PEL) consisting of sets of barriers and wells.
We show that the transition between  thermal  activation (classical) and tunneling (quantum) regimes exhibits universal features of the first and second order phase transition (PT) depending on the PEL for two polariton condensates.
In the presence of dissipation the relative phase of two condensates exhibits non-equilibrium PT from the quantum regime characterized by efficient tunneling of polaritons to the regime of permanent Josephson or Rabi oscillations, where the tunneling is suppressed, respectively.
This analysis paves way for the application of coupled polariton condensates for the realisation of a quantum annealing algorithm.

\pacs{\dots}
\keywords{\dots}

\maketitle

\newcommand{\sn}{\textrm{sn}}
\newcommand{\cn}{\textrm{cn}}
\newcommand{\dn}{\textrm{dn}}
\newcommand{\sd}{\textrm{sd}}
\newcommand{\cd}{\textrm{cd}}
\newcommand{\nd}{\textrm{nd}}
\newcommand{\am}{\textrm{am}}

\section{Introduction \label{sec:introduction}}

@@

Here we specifically focus on the quantum annealing problem that is relevant to the searching algorithm for the global minimum of the potential energy landscape (PEL) consisting of a set of barriers and wells, see e.g.  \cite{Santoro, Das}.
In a purely classical  (thermal) regime the {bosonic quasipartcles cross barriers stochastically} at finite temperature with the help of thermal activation process if the thermal energy is large enough, see Fig. \ref{pic:potential_picture}. Contrary, in a quantum limit the same system undergoes quantum tunneling through the barrier. Obviously, the shape (height and thickness) of the barrier plays an essential role in this case \cite{Das, Lewenstein}. It has been shown recently in \cite{Lewenstein} that the annealing algorithm relies in general case on the combination of thermal annealing and quantum tunneling. The     
collective (bosonic) character of polariton condensation could be used for the acceleration of the physical implementation of the search algorithm \cite{Yan}. The authors of Refs. \cite{Lago1, Lago2} propose polariton graphs as an analog platform for minimizing the XY -- Hamiltonian by exploring the non-equilibrium character of exciton polariton condensates and mimicking various magnetic phases with them.  An original technique of nonresonant spatially  modulated pumping beam  have been used to imprint two-dimensional polariton graphs with different topology and different heights of potential barriers. Importantly, the phase-locking  between arbitrary neighboring vertices might be practically achieved in this case.
%
\begin{figure}[ht]
\center{\includegraphics[width=0.5\linewidth]{pic/potential_picture.eps}}
\caption{Scketch of two tunnel coupled exciton condensates. The two minima of the trapping potential $U$ are situated in the points $+x_0$ and $-x_0$. At thermal equilibrium the interplay between thermal and quantum annealing effects is governed by the blue shift of condensates, effective temperature of exciton-polaritons and the shape of the potential. \label{pic:potential_picture}}
\end{figure} 
%
In the present paper we study the effect of a finite effective temperature on the coupling of quasi-equilibrium exciton-polariton condensates accounting for the competing thremal and quantum annealing effects.

Recently Yongbao Sun et al \cite{Snoke_2017} reported the observation of a quasi-equilibrium low branch exciton polariton condensate in a high Q-factor microcavity characterised by a remarkably long photon lifetime of $135ps$.
Clearly, exciton polaritons with long lifetime are promising candidates for quantum technologies \cite{Demirchyan}.
One of the crucial advantages of exciton-polaritons over cold atoms consists in a perspective of room temperature operation of polariton devices.
In this context, it is very important to reveal the impact of temperature on the physics of interacting polariton condensates.
It should be also noted that the effective temperature of a polariton gas may be introduced in the quasi-equilibrium approximation, which is not necessarily valid in all cases but constitutes an important starting point of any analysis. Further steps would involve a full kinetic modelling for a non-equilibrium polariton gas.

@@

\section{The Quantum Phase Model \label{sec:quantum_phase}}

@@

It is important to note that the last term in Eq. (1) characterizes the familiar XY--model Hamiltonian  considered in
\cite{Lago1, Lago2}.
The second term in (1) introduces an additional part that has no analogy in the XY-Hamiltonian; it is proportional to $\cos(2\phi)$, where $\phi$ is the phase difference between two polariton condensates.
We will show that this term becomes essential if the barrier separating two condensates is sufficiently large.

It is convenient to introduce the pseudo-spin operators in a form
% 
\begin{subequations}
\begin{align}
\hat{S}_x = & s \cos{\phi} - \sin{\phi} \dfrac{d}{d \phi}, \\
\quad \hat{S}_y = & s \sin{\phi} + \cos{\phi} \dfrac{d}{d \phi}, \\
\hat{S}_z = & -i\dfrac{d}{d \phi},
\end{align}
\label{eq:pseudo_spin_diff}
\end{subequations}
%
The operators defined in Eq. (\ref{eq:pseudo_spin_diff}) obey familiar $SU(2)$ algebra commutation relations $[\hat{S}_i, \hat{S}_j] = i\epsilon_{ijk}\hat{S}_k,  i,j,k=x,y,z$.
After some straightforward calculations from (\ref{eq:hamiltonian_spin}) one can obtain
%
\begin{equation}
\begin{array}{lcl}
H & = & -(\alpha - \beta \sin^2 \phi) \dfrac{d^2}{d \phi^2} - (\beta(s-\dfrac{1}{2})\sin{2\phi} - B\sin{\phi}) \dfrac{d}{d \phi} \\
&& - Bs \cos \phi - \beta s^2 \sin^2 \phi - \beta s \cos^2 \phi.
\end{array}
\label{eq:hamiltinian_s}
\end{equation}
%
Thereafter we assume that both condensates are composed by macroscopically large numbers of particles, i.e. the $s = N/2 >> 1$ is C - number.
The general stationary Schroedinger equation with a Hamiltonian (\ref{eq:hamiltinian_s}) writes:
%
\begin{equation}
\begin{array}{l}
(\alpha - \beta \sin^2 \phi) \dfrac{d^2 \Phi}{d \phi^2} + (\beta s\sin{2\phi}-B\sin{\phi})) \dfrac{d \Phi}{d \phi} \\
+ (E + B s \cos \phi + \beta s^2 \sin^2 \phi) \Phi = 0,
\end{array}
\label{eq:schrodinger}
\end{equation}
%
where $\Phi \equiv \Phi(\phi)$ is some $2\pi$-periodic wavefunction that characterizes quantum phase properties \cite{Anglin}.
It is possible to eliminate the term with first derivative in Eq. (\ref{eq:schrodinger}).
In order to find the explicite form of the function $\Phi$, it is convenient to substitute it by:
%
\begin{equation}
\Phi(\phi) = \Psi(z) \exp \Big( s \ln \dn{z} - \dfrac{\Lambda s}{2 \sqrt{\lambda (1 - \lambda)}} \arctan \Big( \sqrt{\dfrac{\lambda}{1 - \lambda}} \cn{z} \Big) \Big),
\label{eq:transformation_of_phi}
\end{equation}
%
where $z = \int \limits_0^\phi \dfrac{d \xi}{\sqrt{1 - \lambda \sin^2 \xi}} = F(\phi, \lambda)$ is a new phase variable; $F(\phi, \lambda)$ is the incomplete elliptic integral of the first kind.
In (\ref{eq:transformation_of_phi}) we have introduced dimension-less parameters $\Lambda = \dfrac{B}{\alpha s}$, $\lambda = \dfrac{\beta}{\alpha}$.
Inserting (\ref{eq:transformation_of_phi}) into (\ref{eq:schrodinger}) we arrive to the familiar form of a Schr\"odinger equation
% 
\begin{equation}
\alpha \frac{d^2\Psi}{dz^2} + (E - V(z))\Psi = 0.
\label{eq:schrodinger_usual}
\end{equation}
%
This equation describes an effective particle with the mass  $m = \dfrac{\hbar^2}{2 \alpha}$ and energy $E$, that is confined by the potential $V(z) = \alpha s^2 V_0(z)$ with
%
\begin{equation}
V_0(z) = \frac{ (\frac{1}{4} \Lambda^2 - \lambda (1 - \lambda)) \sn^2{z} - \Lambda \cn{z}}{\dn^2{z}}.
\label{eq:potential}
\end{equation}
%

The dependences of the trapping potential $V_0(z)$ on the phase variable $z$ are shown in Fig. \ref{pic:phase_potential}. The period of the functions is  $4K(\lambda)$.
The dependence of the $\lambda$ parameter on normalized half of the inter-well distance $\dfrac{x_0}{a}$ can be approximated by $\lambda = 0.5 \Big( \exp \Big[ \dfrac{2 x_0^2}{a^2} \Big] - 1 \Big)^{-1}$ by using variational approach,  see Suppl. Materials.
The well-known behaviour of the quantum phase mesoscopic Josephson junction (that map to the XY--model, \citep{Lago1, Lago2} ) can be recovered from (\ref{eq:hamiltonian_spin}), (\ref{eq:potential}) at $\lambda = 0$ and shown by red curves in Fig. \ref{pic:phase_potential}. This limit corresponds to  infinitely large inter-well distances, with $x_0 \to \infty$.
On the other hand, the $\lambda$ -- parameter exhibits a sharp increase at $\dfrac{x_0}{a} \ll 1$.
Obviously, in this limit our EPJJ model based on the assumption of a relatively weak coupling between trapped condensates breaks down.
Below we are focusing on $\lambda$ -- parameters which belong to the range of $0 < \lambda < 1$ and correspond to the moderate values of $\dfrac{x_0}{a}$ such as $\dfrac{x_0}{a} \geq 0.45$. 
%
\begin{figure}[ht]
\begin{center}\includegraphics[width=1\linewidth]{pic/potentials.eps}
\end{center}
\caption{
Effective dimension-less potential $V_0(z)$ for (a) $\Lambda = 100$, (b) $\Lambda = 0.5$ , and (c) $\Lambda = 0.01$ as a function of the elliptic integral phase coordinate $z$.
Points $\pm z_{min}$ and $\pm z_{max}$ in (b) correspond to global minima and maxima of the potential, respectively; $E$ is the energy of the particle that experiences either quantum tunneling or thermal activation, see Section \ref{sec:quantum_classical}.
\label{pic:phase_potential}}
\end{figure}
%
The analysis of the quantum phase can be performed in three domains determined by vital $\Lambda$--parameter values, \cite{Anglin}.

@@

\section{Annealing problem versus Quantum-classical PT's \label{sec:quantum_classical}}

@@

To reveal the thermodynamic properties of the system it is necessary to consider the small amplitude oscillations at the bottom, $z=0$, of the inverted potential $-V(z)$.
The action in this case reads: 
%
\begin{equation}
S_0 = \Delta V \tau_p (E).
\label{eq:thermal_action}
\end{equation}
%

The expansion of the potential $V(z)$ into a series and excluding the constant value $V(0)$ gives:
%
\begin{equation}
\begin{array}{lcl}
V(z) & = & \alpha s^2 \Big[ \left( \frac{1}{4} \Lambda^2  - \frac{2\lambda - 1}{2} \Lambda - \lambda (1 - \lambda) \right) z^2 \\
&& + \left( \frac{2\lambda - 1}{12} \Lambda^2 - \frac{16\lambda^2 - 16\lambda + 1}{24} \Lambda - \frac{\lambda (1 - \lambda) (2\lambda - 1)}{3} \right) z^4 + o(z^4) \Big].
\end{array}
\label{eq:potential_teylor}
\end{equation}
%
In particular, if the inverted potential $-V(z)$ has the form $z^2 - z^4$ the second order phase transition occurs.
If $-V(z)$ behaves as $z^2 + z^4$ the first order phase transition take place.
The phase boundary between the 1st and 2nd order phase transitions is determined by the relation
%
\begin{equation}
\Lambda = \frac{1 - 16\lambda + 16\lambda^2 + \sqrt{1 + 32\lambda - 32\lambda^2}}{4(2\lambda - 1)}.
\label{eq:order_border}
\end{equation}
%

@@

Let us briefly discuss methods of the experimental study of the quantum-classical PT for the polariton Josephson junctions.
Our estimations show that, for the experimentally accessible polariton interaction strength of $\alpha N=0.6meV$, characteristic of narrow-band semiconductor samples, $T_{0}$ is about $1.2K$ that corresponds to the thermon effective lifetime of $\tau_{0}=3.5ps$.
The value of $T_{0}$ is comparable with a typical temperature of a Bose-Einstein condensation, or, with temperature of the Berezinsky-Kosterlitz-Thouless PT predicted for the dilute weakly interacting exciton polariton gas at thermal equilibrium \cite{Sanvitto,Guillet}.
Some further enhancement of $T_{0}$ may be achieved by the increase of the optical pumping intensity and varying the detuning of exciton and photon modes in a microcavity \cite{Snoke_2017}.

\section{PT's in the presence of dissipation 
\label{sec:non-equilibrium}}

@@

To be more specific in what follows we shall consider the influence of the dissipation on the EPJJ quantum phase properties in the adiabatic limit \cite{Sols}.
We start from GP equations obtained from  (\ref{eq:hamiltonian}) for the mean fields $\psi_{1,2}=<\hat\psi_{1,2}>$ at $\Gamma = 0$ ($B = G$), which read as
%
\begin{equation}
i \hbar \dot{\psi}_{1,2} = -i \kappa \psi_{1,2} + (A|\psi_{1,2}|^2 + 2C |\psi_{2,1}|^2) \psi_{1,2} - (B/2 - C \psi_{1,2}^* \psi_{2,1}) \psi_{2,1}. 
\label{eq:GP}
\end{equation}
%
In (\ref{eq:GP}) we have introduced the dissipation term with $\kappa \simeq \hbar/\tau_{pol}$.
Defining new variables $\Psi_{1,2}$ as $\psi_{1,2} = \Psi_{1,2} \exp(-\kappa t / \hbar)$ for the mean field pseudo-spin components (\ref{eq:pseudo_spin}).
We obtain from (\ref{eq:GP})
% 
\begin{equation}
\begin{array}{lcl}
\dot{S}_x & = & -2 \alpha'(t) S_z S_y; \\
\dot{S}_y & = & 2(\alpha'(t) - \beta'(t)) S_z S_x + S_z. \\
\dot{S}_z & = &  2 \beta'(t) S_x S_y - S_y;
\end{array}
\label{eq:pseudo-spin}
\end{equation}
%
Here we have introduced the dimensionless time $t' = t B / \hbar$.

@@

In the presence of dissipation (solid black curve in Fig. \ref{pic:phase}a) the phase of EPJJ starts from one of the potential minima.
We assume here that the temperature of the system is not sufficient for thermal activation and the quantum tunneling is possible. The system evolves adiabatically and then  crosses the critical value $\beta = 1/2$
due to the decrease of the total number of particles caused by the radiative decay.
The final (temporal) state represents the Josephson or, Rabi oscillation regime.
%
\begin{figure}[ht]
\center{\includegraphics[width=\linewidth]{pic/phase_potential_dissipation.eps}}
\caption{(a) -- Normalized phase difference as a function of time; (b) -- effective potential vs phase parameter $z$ for red, black and blue curves, respectively, and (c) -- reduced thermon period $\tau_{p} / \tau_0$ ($\tau_0 = \pi \hbar / \alpha s $) vs time for $\kappa/\hbar = 0.1 THz$. 
Initial conditions are $S_z(0) = S_x(0) = 0$, $S_y(0) = 1$, and $\phi(0) = \pi / 2$, respectively.
\label{pic:phase}}
\end{figure}
%

@@

\section{Conclusion \label{sec:conclusion}}

@@

First, it is quantum information processing and especially the quantum annealing problem  with exciton polariton condensates that allow for acceleration of the quantum annealing due to the final state bosonic stimulation {\cite{Yan}.
Notably the results obtained by us can be implemented here in new way.
In order, quantum tunneling effect might be utilized for design exciton polariton phase qubits similarly to superconductor devices, but in optical wavelength domain, \cite{Makhlin}.
Actually, two minima located at the points $-z_{min}$ and $+z_{max}$ of $W$-shape quantum phase potential in Fig. \ref{pic:phase_potential}b can be used for initialization of exciton polariton qubit states $\ket{0}$ and $\ket{1}$ similarly to superconductor flux qubits, respectively.
However, in our case such qubits may be tailored by external optical or electrical pump.
As a result, quantum annealing problem can be experimentally solved with help of this type of qubits \cite{Johnson}.

Another promising area of potential applications of the obtained results is quantum metrology.
In particular, EPJJs can be useful for realisation of interferometers and/or gyroscopes operating with the phase of  exciton polariton condensates in the quantum regime \cite{Pezze, Gulevich}.
The high accuracy measurements of the phase relations between polariton condensates would be needed in this case.
It is possible to show that for the Fock regime the dispersion of the quantum phase approaches the so-called Heisenberg limit ${<(\Delta\phi)^2> \propto \dfrac{1}{N^2}}$ that is the minimal accessible error for the phase measurement for the given total number of particles.

\section{Acknowledgement
\label{sec:acknowledgement}}
We acknowledge the financial support from RFBR, Grants No. 15-52-52001 and No. 14-02-97503.
A. Kavokin acknowledges the support from HORIZON 2020 RISE project CoExAn (Grant No. 644076) and the EPSRC Established Career fellowship on Quantum Polaritonics RP 008833.

>>>>

\documentclass[fleqn,10pt]{wlscirep}

@@

@@

% Elliptic functions
\newcommand{\sn}{\textrm{sn}}
\newcommand{\cn}{\textrm{cn}}
\newcommand{\dn}{\textrm{dn}}
\newcommand{\sd}{\textrm{sd}}
\newcommand{\cd}{\textrm{cd}}
\newcommand{\nd}{\textrm{nd}}
\newcommand{\am}{\textrm{am}}

% Title and authors
\title{Exciton-polariton Josephson junctions at finite temperatures}

\author[1]{M. E. Lebedev}
\author[1]{D. A. Dolinina}
\author[2]{Kuo-Bin Hong}
\author[2]{Tien-Chang Lu}
\author[3, 4, 5]{A. V. Kavokin}
\author[1, 6, *]{A. P. Alodjants}

\affil[1]{ITMO University, St. Petersburg 197101, Russia}
\affil[2]{Department of Photonics, National Chiao Tung University, Hsinchu 300, Taiwan}
\affil[3]{Spin Optics Laboratory, St. Petersburg State University, Ul’anovskaya, Peterhof, St. Petersburg 198504, Russia}
\affil[4]{School of Physics and Astronomy, University of Southampton, SO17 1BJ Southampton, United Kingdom}
\affil[5]{Istituto CNR-SPIN, Viale del Politecnico 1, I-00133, Rome, Italy}
\affil[6]{Vladimir State University named after A. G. and N. G. Stoletovs, Gorkii Street 87, Vladimir, Russia}

\affil[*]{alexander\_ap@list.ru}

% \keywords{\dots}

\begin{abstract}

We consider finite temperature effects in a non-standard Bose-Hubbard model for an exciton polariton Josephson junction (EPJJ) that is characterised by complicated potential energy landscapes (PEL) consisting of sets of barriers and wells.
We show that the transition between  thermal  activation (classical) and tunneling (quantum) regimes exhibits universal features of the first and second order phase transition (PT) depending on the PEL for two polariton condensates.
In the presence of dissipation the relative phase of two condensates exhibits non-equilibrium PT from the quantum regime characterized by efficient tunneling of polaritons to the regime of permanent Josephson or Rabi oscillations, where the tunneling is suppressed, respectively.
This analysis paves way for the application of coupled polariton condensates for the realisation of a quantum annealing algorithm.

@@

\flushbottom
\maketitle

% (?)
\thispagestyle{empty}

\section*{Introduction}

@@

Here we specifically focus on the quantum annealing problem that is relevant to the searching algorithm for the global minimum of the potential energy landscape (PEL) consisting of a set of barriers and wells, see e.g.  \cite{Santoro, Das}.
In a purely classical  (thermal) regime the {bosonic quasipartcles cross barriers stochastically} at finite temperature with the help of thermal activation process if the thermal energy is large enough, see Fig. \ref{pic:potential_picture}. Contrary, in a quantum limit the same system undergoes quantum tunneling through the barrier. Obviously, the shape (height and thickness) of the barrier plays an essential role in this case \cite{Das, Lewenstein}. It has been shown recently in \cite{Lewenstein} that the annealing algorithm relies in general case on the combination of thermal annealing and quantum tunneling. The     
collective (bosonic) character of polariton condensation could be used for the acceleration of the physical implementation of the search algorithm \cite{Yan}. The authors of Refs. \cite{Lago1, Lago2} propose polariton graphs as an analog platform for minimizing the XY -- Hamiltonian by exploring the non-equilibrium character of exciton polariton condensates and mimicking various magnetic phases with them.  An original technique of nonresonant spatially  modulated pumping beam  have been used to imprint two-dimensional polariton graphs with different topology and different heights of potential barriers. Importantly, the phase-locking  between arbitrary neighboring vertices might be practically achieved in this case.
%
\begin{figure}[ht]
\center{\includegraphics[width=0.5\linewidth]{pic/potential_picture.eps}}
\caption{Scketch of two tunnel coupled exciton condensates. The two minima of the trapping potential $U$ are situated in the points $+x_0$ and $-x_0$. At thermal equilibrium the interplay between thermal and quantum annealing effects is governed by the blue shift of condensates, effective temperature of exciton-polaritons and the shape of the potential. \label{pic:potential_picture}}
\end{figure} 
%

In the present paper we study the effect of a finite effective temperature on the coupling of quasi-equilibrium exciton-polariton condensates accounting for the competing thremal and quantum annealing effects.  
Recently Yongbao Sun et al \cite{Snoke_2017} reported the observation of a quasi-equilibrium low branch exciton polariton condensate in a high Q-factor microcavity characterised by a remarkably long photon lifetime of $135ps$.
Clearly, exciton polaritons with long lifetime are promising candidates for quantum technologies \cite{Demirchyan}.
One of the crucial advantages of exciton-polaritons over cold atoms consists in a perspective of room temperature operation of polariton devices.
In this context, it is very important to reveal the impact of temperature on the physics of interacting polariton condensates.
It should be also noted that the effective temperature of a polariton gas may be introduced in the quasi-equilibrium approximation, which is not necessarily valid in all cases but constitutes an important starting point of any analysis. Further steps would involve a full kinetic modelling for a non-equilibrium polariton gas.

@@

\section*{Results}

\subsection*{The Quantum Phase Model \label{sec:quantum_phase}}

@@

It is important to note that the last term in Eq. (1) characterizes the familiar XY--model Hamiltonian  considered in
\cite{Lago1, Lago2}.
The second term in (1) introduces an additional part that has no analogy in the XY-Hamiltonian; it is proportional to $\cos(2\phi)$, where $\phi$ is the phase difference between two polariton condensates.
We will show that this term becomes essential if the barrier separating two condensates is sufficiently large.
It is convenient to introduce the pseudo-spin operators in a form
% 
\begin{subequations}
\begin{align}
\hat{S}_x = & s \cos{\phi} - \sin{\phi} \dfrac{d}{d \phi}, \\
\quad \hat{S}_y = & s \sin{\phi} + \cos{\phi} \dfrac{d}{d \phi}, \\
\hat{S}_z = & -i\dfrac{d}{d \phi},
\end{align}
\label{eq:pseudo_spin_diff}
\end{subequations}
%
The operators defined in Eq. (\ref{eq:pseudo_spin_diff}) obey familiar $SU(2)$ algebra commutation relations $[\hat{S}_i, \hat{S}_j] = i\epsilon_{ijk}\hat{S}_k,  i,j,k=x,y,z$.
After some straightforward calculations from (\ref{eq:hamiltonian_spin}) one can obtain
%
\begin{equation}
\begin{array}{lcl}
H & = & -(\alpha - \beta \sin^2 \phi) \dfrac{d^2}{d \phi^2} - (\beta(s-\dfrac{1}{2})\sin{2\phi} - B\sin{\phi}) \dfrac{d}{d \phi} \\
&& - Bs \cos \phi - \beta s^2 \sin^2 \phi - \beta s \cos^2 \phi.
\end{array}
\label{eq:hamiltinian_s}
\end{equation}
%
Thereafter we assume that both condensates are composed by macroscopically large numbers of particles, i.e. the $s = N/2 >> 1$ is C - number.
The general stationary Schroedinger equation with a Hamiltonian (\ref{eq:hamiltinian_s}) writes:
%
\begin{equation}
\begin{array}{l}
(\alpha - \beta \sin^2 \phi) \dfrac{d^2 \Phi}{d \phi^2} + (\beta s\sin{2\phi}-B\sin{\phi})) \dfrac{d \Phi}{d \phi} \\
+ (E + B s \cos \phi + \beta s^2 \sin^2 \phi) \Phi = 0,
\end{array}
\label{eq:schrodinger}
\end{equation}
%
where $\Phi \equiv \Phi(\phi)$ is some $2\pi$-periodic wavefunction that characterizes quantum phase properties \cite{Anglin}.
It is possible to eliminate the term with first derivative in Eq. (\ref{eq:schrodinger}).
In order to find the explicite form of the function $\Phi$, it is convenient to substitute it by:
%
\begin{equation}
\Phi(\phi) = \Psi(z) \exp \Big( s \ln \dn{z} - \dfrac{\Lambda s}{2 \sqrt{\lambda (1 - \lambda)}} \arctan \Big( \sqrt{\dfrac{\lambda}{1 - \lambda}} \cn{z} \Big) \Big),
\label{eq:transformation_of_phi}
\end{equation}
%
where $z = \int \limits_0^\phi \dfrac{d \xi}{\sqrt{1 - \lambda \sin^2 \xi}} = F(\phi, \lambda)$ is a new phase variable; $F(\phi, \lambda)$ is the incomplete elliptic integral of the first kind.
In (\ref{eq:transformation_of_phi}) we have introduced dimension-less parameters $\Lambda = \dfrac{B}{\alpha s}$, $\lambda = \dfrac{\beta}{\alpha}$.
Inserting (\ref{eq:transformation_of_phi}) into (\ref{eq:schrodinger}) we arrive to the familiar form of a Schr\"odinger equation
% 
\begin{equation}
\alpha \frac{d^2\Psi}{dz^2} + (E - V(z))\Psi = 0.
\label{eq:schrodinger_usual}
\end{equation}
%
This equation describes an effective particle with the mass  $m = \dfrac{\hbar^2}{2 \alpha}$ and energy $E$, that is confined by the potential $V(z) = \alpha s^2 V_0(z)$ with
%
\begin{equation}
V_0(z) = \frac{ (\frac{1}{4} \Lambda^2 - \lambda (1 - \lambda)) \sn^2{z} - \Lambda \cn{z}}{\dn^2{z}}.
\label{eq:potential}
\end{equation}
%

The dependences of the trapping potential $V_0(z)$ on the phase variable $z$ are shown in Fig. \ref{pic:phase_potential}. The period of the functions is  $4K(\lambda)$.
The dependence of the $\lambda$ parameter on normalized half of the inter-well distance $\dfrac{x_0}{a}$ can be approximated by $\lambda = 0.5 \Big( \exp \Big[ \dfrac{2 x_0^2}{a^2} \Big] - 1 \Big)^{-1}$ by using variational approach,  see Suppl. Materials.
The well-known behaviour of the quantum phase mesoscopic Josephson junction (that map to the XY--model, \cite{Lago1, Lago2} ) can be recovered from (\ref{eq:hamiltonian_spin}), (\ref{eq:potential}) at $\lambda = 0$ and shown by red curves in Fig. \ref{pic:phase_potential}. This limit corresponds to  infinitely large inter-well distances, with $x_0 \to \infty$.
On the other hand, the $\lambda$ -- parameter exhibits a sharp increase at $\dfrac{x_0}{a} \ll 1$.
Obviously, in this limit our EPJJ model based on the assumption of a relatively weak coupling between trapped condensates breaks down.
Below we are focusing on $\lambda$ -- parameters which belong to the range of $0 < \lambda < 1$ and correspond to the moderate values of $\dfrac{x_0}{a}$ such as $\dfrac{x_0}{a} \geq 0.45$. 
%
\begin{figure}[ht]
\begin{center}\includegraphics[width=1\linewidth]{pic/potentials.eps}
\end{center}
\caption{
Effective dimension-less potential $V_0(z)$ for (a) $\Lambda = 100$, (b) $\Lambda = 0.5$ , and (c) $\Lambda = 0.01$ as a function of the elliptic integral phase coordinate $z$.
Points $\pm z_{min}$ and $\pm z_{max}$ in (b) correspond to global minima and maxima of the potential, respectively; $E$ is the energy of the particle that experiences either quantum tunneling or thermal activation, see Section \ref{sec:quantum_classical}.
\label{pic:phase_potential}}
\end{figure}
%
The analysis of the quantum phase can be performed in three domains determined by vital $\Lambda$--parameter values, \cite{Anglin}.

@@

\subsection*{Annealing problem versus Quantum-classical PT's \label{sec:quantum_classical}}

@@

To reveal the thermodynamic properties of the system it is necessary to consider the small amplitude oscillations at the bottom, $z=0$, of the inverted potential $-V(z)$.
The action in this case reads: 
%
\begin{equation}
S_0 = \Delta V \tau_p (E).
\label{eq:thermal_action}
\end{equation}
%
The expansion of the potential $V(z)$ into a series and excluding the constant value $V(0)$ gives:
%
\begin{equation}
\begin{array}{lcl}
V(z) & = & \alpha s^2 \Big[ \left( \frac{1}{4} \Lambda^2  - \frac{2\lambda - 1}{2} \Lambda - \lambda (1 - \lambda) \right) z^2 \\
&& + \left( \frac{2\lambda - 1}{12} \Lambda^2 - \frac{16\lambda^2 - 16\lambda + 1}{24} \Lambda - \frac{\lambda (1 - \lambda) (2\lambda - 1)}{3} \right) z^4 + o(z^4) \Big].
\end{array}
\label{eq:potential_teylor}
\end{equation}
%
In particular, if the inverted potential $-V(z)$ has the form $z^2 - z^4$ the second order phase transition occurs.
If $-V(z)$ behaves as $z^2 + z^4$ the first order phase transition take place.
The phase boundary between the 1st and 2nd order phase transitions is determined by the relation
%
\begin{equation}
\Lambda = \frac{1 - 16\lambda + 16\lambda^2 + \sqrt{1 + 32\lambda - 32\lambda^2}}{4(2\lambda - 1)}.
\label{eq:order_border}
\end{equation}
%

@@

Let us briefly discuss methods of the experimental study of the quantum-classical PT for the polariton Josephson junctions.
Our estimations show that, for the experimentally accessible polariton interaction strength of $\alpha N=0.6meV$, characteristic of narrow-band semiconductor samples, $T_{0}$ is about $1.1K$ that corresponds to the thermon effective lifetime of $\tau_{0}\approx 7 ps$.
The value of $T_{0}$ is comparable with a typical temperature of a Bose-Einstein condensation, or, with temperature of the Berezinsky-Kosterlitz-Thouless PT predicted for the dilute weakly interacting exciton polariton gas at thermal equilibrium \cite{Sanvitto,Guillet}.
Some further enhancement of $T_{0}$ may be achieved by the increase of the optical pumping intensity and varying the detuning of exciton and photon modes in a microcavity \cite{Snoke_2017}.

\subsection*{PT's in the presence of dissipation\label{sec:non-equilibrium}}

@@

To be more specific in what follows we shall consider the influence of the dissipation on the EPJJ quantum phase properties in the adiabatic limit \cite{Sols}.
We start from GP equations obtained from  (see supplementary equation (S5)) for the mean fields $\psi_{1,2}=<\hat\psi_{1,2}>$ at $\Gamma = 0$ ($B = G$), which read as
%
\begin{equation}
i \hbar \dot{\psi}_{1,2} = -i \kappa \psi_{1,2} + (A|\psi_{1,2}|^2 + 2C |\psi_{2,1}|^2) \psi_{1,2} - (B/2 - C \psi_{1,2}^* \psi_{2,1}) \psi_{2,1}. 
\label{eq:GP}
\end{equation}
%
In (\ref{eq:GP}) we have introduced the dissipation term with $\kappa \simeq \hbar/\tau_{pol}$.
Defining new variables $\Psi_{1,2}$ as $\psi_{1,2} = \Psi_{1,2} \exp(-\kappa t / \hbar)$ for the mean field pseudo-spin components (see supplementary equation (S6)).
We obtain from (\ref{eq:GP})
% 
\begin{equation}
\begin{array}{lcl}
\dot{S}_x & = & -2 \alpha'(t) S_z S_y; \\
\dot{S}_y & = & 2(\alpha'(t) - \beta'(t)) S_z S_x + S_z. \\
\dot{S}_z & = &  2 \beta'(t) S_x S_y - S_y;
\end{array}
\label{eq:pseudo-spin}
\end{equation}
%
Here we have introduced the dimensionless time $t' = t B / \hbar$.

@@

In the presence of dissipation (solid black curve in Fig. \ref{pic:phase}a) the phase of EPJJ starts from one of the potential minima.
We assume here that the temperature of the system is not sufficient for thermal activation and the quantum tunneling is possible. The system evolves adiabatically and then  crosses the critical value $\beta = 1/2$
due to the decrease of the total number of particles caused by the radiative decay.
The final (temporal) state represents the Josephson or, Rabi oscillation regime.
%
\begin{figure}[ht]
\center{\includegraphics[width=\linewidth]{pic/phase_potential_dissipation.eps}}
\caption{(a) -- Normalized phase difference as a function of time; (b) -- effective potential vs phase parameter $z$ for red, black and blue curves, respectively, and (c) -- reduced thermon period $\tau_{p} / \tau_0$  vs time for $\kappa/\hbar = 0.1 THz$. 
Initial conditions are $S_z(0) = S_x(0) = 0$, $S_y(0) = 1$, and $\phi(0) = \pi / 2$, respectively.
\label{pic:phase}}
\end{figure}
%

@@

\section*{Discussion \label{sec:conclusion}}

@@

In order, it is quantum information processing and especially the quantum annealing problem  with exciton polariton condensates that allow for acceleration of the quantum annealing due to the final state bosonic stimulation {\cite{Yan}.
Notably the results obtained by us can be implemented here in new way.
In order, quantum tunneling effect might be utilized for design exciton polariton phase qubits similarly to superconductor devices, but in optical wavelength domain, \cite{Makhlin}.
Actually, two minima located at the points $-z_{min}$ and $+z_{max}$ of $W$-shape quantum phase potential in Fig. \ref{pic:phase_potential}b can be used for initialization of exciton polariton qubit states $\ket{0}$ and $\ket{1}$ similarly to superconductor flux qubits, respectively.
However, in our case such qubits may be tailored by external optical or electrical pump.
As a result, quantum annealing problem can be experimentally solved with help of this type of qubits \cite{Johnson}.

\section*{Methods}
\dots
